Variation on a Variational Principle

### 2020-09-22 draft1.tex
done proof of main theorem. many TODO eg feasible flow thm

\begin{lemma}
%Let $P_n$ be an ideal pyramid whose base is an $n$-gon and suppose,
%\begin{itemize}
%\item each boundary edge of the base face is assigned a dihedral angle $\alpha_i$ such that for adjacent edges, $\alpha_i + \alpha_{i+1} < \pi$;
%\item we are given a decomposition of the base face into triangles by adding new edges. 
%\end{itemize}
Let $P_n$ be a base-angled ideal pyramid, and suppose we are given a
decomposition of the base face into triangles by adding new edges.  One gets an
obvious corresponding triangulation of $P_n$, where a new face is added for each
new edge. Then there is an assignment of a dihedral angle to each edge of each
ideal tetrahedron in this triangulation such that
\begin{itemize}
\item each tetrahedron is an angled ideal tetrahedron;
\item the sum of dihedral angles around each new edge is $\pi$;
\item the dihedral angles of the edges of the original base face are the same as
	before.
\end{itemize} 
\end{lemma}

@@

\begin{figure}
\label{f:ideal_pyramid_arg}
\includegraphics[height=8cm]{more_pictures/angle_split.png}
\end{figure}
Suppose the edges are labelled $e_i$,
which goes between vertices $v_i$ and $v_{i+1}$,
and suppose $e_i$ is assigned dihedral angle $\alpha_i$.
Let $e'$ be a new edge addeed to the base face of $P_n$
such that it separates the base face into a triangle and
an $(n-1)$-gon;
suppose the sides of the triangle are
$e_i, e_{i+1}$, and $e'$.
The new face corresponding to $e'$ separates $P_n$ into
an ideal tetrahedron $T$ and an ideal pyramid $P_{n-1}$.
We assign the dihedral angle of $\pi - \alpha_i - \alpha_{i+1}$
to $e'$ in $T$, and assign $\alpha_i + \alpha_{i+1}$ to $e'$ in $P_{n-1}$.
Clearly the sum of dihedral angles condition is satisfied
in $T$ and $P_{n-1}$.
It remains to check that the dihedral angles assigned to the vertical edges
are correct.
For the vertical edge associated to $v_j$ for $j \neq i, i+2$,
there is nothing to check;
for $j = i$, the dihedral angles are
$\pi - \alpha_i - (\pi - \alpha_i - \alpha_{i-1})$
in $T$ and $\pi - \alpha_{i-1} - (\alpha_i + \alpha_{i+1})$ in $P_{n-1}$,
which sum to $\pi - \alpha_i - \alpha_{i+1}$;
it is similar for $j = i+2$.

@@

\comment{
\begin{define}
Given a face $f$ of the original link diagram the \emph{augmentation sequence}
of $f$, denoted $\sigma_f$, is a cyclical sequence of symbols $0, L, R$ indexed
by vertices of $f$ going counter-clockwise, where, for $v \in \partial f$, the
$v$-th element is
\[
\sigma_f^v = 
\begin{cases}
0 \text{ if $v$ is not augmented} \\
L \text{ if $v$ is left-augmented} \\
R \text{ if $v$ is right-augmented}
\end{cases}
\]
\end{define}
}

\comment{

\begin{prop}
Suppose $L$ is obtained from augmenting $K$,
so that for each face $f$ of the link projection of $K$,
the augmentation sequence $\sigma_f$ has no $L$ adjacent to $R$.
Then the TODO CONTINUE HERE
\end{prop}
}

\comment{
\begin{prop}
Let $\Sigma$ be a closed surface with a cellular decomposition.
Suppose we have labelled some edges by either $`+'$ or $`-'$,
such that for each face $f$,
\begin{itemize}
	\item no two adjacent edges have the same label,
	\item the length of consecutive labelled edges is at least 2.
\end{itemize}
Then one can further decompose faces along new edges,
and extend the labelling of edges,
such that every vertex is adjacent to the same number of
$`+'$'s and $`-'$'s.
\end{prop}

\begin{proof}
Consider a face $f$ and consider a maximal contiguous sequence
of labelled edges, $e_1,\ldots,e_k$ in $\del f$.
they span the vertices $v_0,\ldots,v_k$.
Suppose $e_1$ is labelled $`+'$ (the other case is similar).
Then $e_2$ is labelled $`-'$, $e_3$ is $`+'$, and so on.
\end{proof}

\textbf{Case 1}: $k$ is odd. Then we cut $f$ along a new edge
$\overline{v_0 v_k}$, and label this edge by $`-'$.

\textbf{Case 2}: $k = 2$ and . Then we cut $f$ along a new edge
\end{proof}

}

%\begin{theorem}
%Let $K$ be a weakly prime, alternating link with diagram $D$ with no bigons. Let
%$L$ be a link obtained from augmenting $K$ such that there exist no sequence of
%augmentations where a Left augmentation is adjacent to a Right augmentation.
%Then $L$ is hyperbolic.
%\end{theorem}
\begin{theorem}
Let $K$ be a weakly prime, alternating link whose diagram has no bigons.
Let $L$ be a link obtained from augmenting $K$.
%such that
%no edge spans left- and right-augmented vertices.
Then $L$ is hyperbolic.
\end{theorem}

@@

By \prpref{p:tori_decomp}, $\toruscomp{L}$ can be
obtained by gluing two torihedra $\sT_T(L),\sT_B(L)$
with graphs $\Gamma_T(L),\Gamma_B(L)$.
We make them degenerate angled torihedra by assigning $\theta^*$'s
to edges of the bow-ties as in \figref{f:bowtie_angles},
and assign $\pi/2$ to all other edges.

@@

Furthermore, we can obtain an angle-splitting of $\sT_T(L)$
(and similarly $\sT_B(L)$) by modifying the angle-splitting
for $\sT_T(K)$.

Hmm

\begin{proof}
By \cite[Theorem 7.5]{CKP2},
$\toruscomp{K}$ can be decomposed
into two torihedra whose graphs are the projection graph of $K$,
which we denote by $\Gamma$.
By assigning angles $\pi/2$ to each edge of the torihedra graphs,
we can invoke \cite[Theorem 3]{BandS} to give us a circle pattern
on the projection graph.

By \prpref{p:tori_decomp}, $\toruscomp{L}$ can be
obtained by gluing two torihedra with graph $\Gamma_T(L)$ and $\Gamma_B(L)$.
Recall that $\Gamma_T(L)$ is obtained from $\Gamma$ by some slight
modifications, specifically by replacing augmented vertices with bow-ties.
We assign angles to edges of the bow-ties as in
\figref{f:bowtie_angles},
and leave untouched edges with the same assignment of $\pi/2$;
we do the same for $\Gamma_B(L)$.

%\begin{figure}
%\label{f:bowtie_angles}
%\includegraphics[width=5cm]{more_pictures/horizontal_bowtie.png}
%\end{figure}

(Hmmm... might be misleading, I think $\Psi$ should be a cellular map
$\Gamma_T(L) \to \Gamma_T(K)$, but the latter happens to be
same as $\Gamma$, so that's why it's confusing!)
We can think of a bow-tie in $\Gamma_T(L)$ as ``thickened edges" in the
following sense: consider the bow-tie at a crossing $v$,
and let the edges $e,e'$ of $\Gamma$ at $v$
be associated to the over-strand of $K$.
There is a cellular map from each triangle of the bow-tie
on to $e, e'$ that identifies the $\pi/2$- and 0-labelled edges,
and collapses the $\pi$-labelled edges onto $v$
(TODO see same figure).
It is not hard to see that this map extends to a cellular map
$\Psi: \Gamma_T(L) \to \Gamma$,
and in fact extends to the faces as well.
A similar story applies to $\Gamma_B$.

Now we construct an extended circle pattern for $\Gamma_T(L)$.
We will give a more precise description,
but let us first give the picture.
Under $\Psi$,
each face $f$ of $\Gamma_T(L)$ that is not in a bow-tie
is identified with a face $\bar{f}$ of $\Gamma$,
so we assign the circle of $\bar{f}$ to $f$;
each triangle face $f$ in a bow-tie is collapsed to an edge,
and we assign to $f$ the circle circumscribing the
face of $\Gamma$ that meets $f$ along the 0-labelled edge.
This is a ``degenerate" circle pattern that can be thought of as
the limit of some (singular) circle patterns.

Let us make this more precise,
describing an extended circle pattern $c$
on $\Gamma_T(L)$.
Let the circle pattern for $\Gamma$ be given by
$\bar{c} = ((r_f),(\vphi_{\vec{e}}))$,
where recall $r_f$ is the radius of the circle $C_f$
circumscribing $f$ and $\vphi_{\vec{e}}$
is half the angle subtended by the edge $e$
at the center of $C_f$.
For non-bow-tie faces $f$ of $\Gamma_T(L)$,
$\Psi$ identifies $f$ with the face $\bar{f}$ of $\Gamma$.
We set $r_f(c) = r_{\bar{f}}(\bar{c})$.
For $\vphi$ of its edges,
if $\vec{e} \in \del f$ is labelled $\pi$,
we set $\vphi_{\vec{e}} = 0$;
otherwise, $\Psi$ identifies $\vec{e}$
with the edge $\vec{e}'$ in $\Gamma$,
then we set $\vphi_{\vec{e}}(c) = \vphi_{\vec{e}'}(\bar{c})$.

For a triangular face $f$ of a bow tie,
let $\vec{e}_0, \vec{e}_{\pi/2}, \vec{e}_{\pi}$
be the edges of $f$ labelled $0,\pi/2, \pi$ respectively.
We set $r_f(c) = r_{f'}(c)$,
where $f'$ is the face adjacent to $f$ across $e_0$.
Then we set
$\vphi_{\vec{e}_{\pi/2}} = \vphi_{\cev{e}_0} = \pi - \vphi_{\vec{e}_0}$
and $\vphi_{\vec{e}_{\pi}} = 0$.

It can be easily checked that $c$ satisfies the conditions of
\prpref{p:nghd_lift}.
We would like to apply \lemref{l:circpattern_polyhedra}
to obtain a polyhedral decomposition of the torihedra,
but there are edges with $\vphi_{\vec{e}}$
(specifically, those with $\theta_e = \pi$).
%We want a deformation of $\theta_{\bullet}$ of $c$
%that remains vertex non-singular,
%and gives us an honest ?? circle pattern.
Hence, we want to find a vector $a = \sum a^e \ddd{\theta_e}$
in the tangent space of $\Theta(c)$
that maintains vertex non-singularity,
i.e.  such that $\sum_{e \ni v} a^e = 0$ for all vertices $v$,
and for edges $e$ with $\theta_e = \pi$,
we have $a^e < 0$.
By \prpref{p:nghd_lift}, there exists $\veps > 0$
such that the path $\gamma(t) = \Theta(c) + t \cdot a$, $t\in [0,\veps]$,
in $\TTT$ can be lifted to a path $\tilde{\gamma}$ in $\CCC$.

To find such $a$, we need to modify $\Gamma_T(L)$ by
cutting faces along new edges.
Consider a face $f$ that is not in a bow-tie.
Suppose the corresponding face $\bar{f}$ of $\Gamma$
had vertices $v_1,\ldots,v_n$ in counter-clockwise order.
Suppose it is the case that if a vertex $v_i$ is left-augmented,
then the augmentation circle intersects $\bar{f}$
(everything is similar if it is right-augmented circles that intersect
$\bar{f}$).
Vertex $v_i$ corresponds to one edge $e_i$ of $\Gamma_T(L)$
if $v_i$ is not augmented or right-augmented,
and corresponds to two edges $e_{i,0}, e_{i,\pi}$
of $\Gamma_T(L)$ if $v_i$ is left-augmented
($e_{i,0}, e_{i,\pi}$ are edges of a single bow-tie,
and $e_{i,0}$ has $\theta_{e_{i,0}} = 0$
and $e_{i,\pi}$ has $\theta_{e_{i,\pi}} = \pi$).

Suppose, after cyclically reindexing, $v_1,\ldots,v_k$
is a maximally contiguous subsequence of left-augmented vertices
of $G(K)$ around the face $\bar{f}$;
the edges around $f$ would start
$e_{1,0}, e_{1,\pi}, e_{2,0}, e_{2,\pi}, \ldots$.
We add new edges across $f$ as follows.

@@

Next suppose there is only one such maximal contiguous subsequence.
If $k = 1$, we add an edge that goes across
$e_{1,0},e_{1,\pi},e_2$
(in the sense that the new edge separates the edges of $f$ into two sets
one of them being those three edges;
since $n\geq 3$, this edge is new).
If $k = 2$, we add edge across $e_{1,0},e_{1,\pi}$
and another edge across $e_{2,0},e_{2,\pi}$
(these two edges do not form a bigon because we've ruled out $k=n$).
If $k \geq 3$,
we add an edge across $e_{1,0},e_{1,\pi},e_{2,0}$
and another edge across $e_{2,\pi},e_{3,0},\ldots,e_{k,\pi}$
(again these two edges do not form a bigon).

@@

See \figref{f:adding_edges}

\begin{figure}
\label{f:adding_edges}
\includegraphics[width=5cm]{more_pictures/one_edge.png}
\includegraphics[width=5cm]{more_pictures/two_edge.png}
\includegraphics[width=5cm]{more_pictures/two_edge_many.png}
\end{figure}

TODO CONTINUE HERE do the $+$ and $-$ on the edges.

Our torihedral decomposition has graph and on augmentation we will give 0 and pi angles. Next we add another edge and give blah angle (we have three cases)
do three cases with pictures

part 3: Use proposition to get circle pattern to get polyhedral decomposition of the link

part 4: get triangulation.

@@

TODO consistent torus symbol, consistent thickened torus symbols

>>>>

\begin{lemma}
\label{l:pyramid_decomp}
%Let $P_n$ be an ideal pyramid whose base is an $n$-gon and suppose,
%\begin{itemize}
%\item each boundary edge of the base face is assigned a dihedral angle $\alpha_i$ such that for adjacent edges, $\alpha_i + \alpha_{i+1} < \pi$;
%\item we are given a decomposition of the base face into triangles by adding new edges. 
%\end{itemize}
Let $P_n$ be a base-angled ideal pyramid, and suppose we are given a
decomposition of the base face into triangles by adding new edges.  One gets an
obvious corresponding triangulation of $P_n$, where a new face is added for each
new edge. Then there is an assignment of a dihedral angle to each edge of each
ideal tetrahedron in this triangulation such that
\begin{itemize}
\item each tetrahedron is an angled ideal tetrahedron;
\item the sum of dihedral angles around each new edge is $\pi$;
\item the dihedral angles of the edges of the original base face are the same as
	before.
\end{itemize} 
\end{lemma}

@@

\begin{figure}
\label{f:ideal_pyramid_arg}
\includegraphics[height=8cm]{more_pictures/angle_split.png}
\end{figure}
Suppose the edges are labeled $e_i$,
which goes between vertices $v_i$ and $v_{i+1}$,
and suppose $e_i$ is assigned dihedral angle $\alpha_i$.
Let $e'$ be a new edge addeed to the base face of $P_n$
such that it separates the base face into a triangle and
an $(n-1)$-gon;
suppose the sides of the triangle are
$e_i, e_{i+1}$, and $e'$.
The new face corresponding to $e'$ separates $P_n$ into
an ideal tetrahedron $T$ and an ideal pyramid $P_{n-1}$.
We assign the dihedral angle of $\pi - \alpha_i - \alpha_{i+1}$
to $e'$ in $T$, and assign $\alpha_i + \alpha_{i+1}$ to $e'$ in $P_{n-1}$.
Clearly the sum of dihedral angles condition is satisfied
in $T$ and $P_{n-1}$.
It remains to check that the dihedral angles assigned to the vertical edges
are correct.
For the vertical edge associated to $v_j$ for $j \neq i, i+2$,
there is nothing to check;
for $j = i$, the dihedral angles are
$\pi - \alpha_i - (\pi - \alpha_i - \alpha_{i-1})$
in $T$ and $\pi - \alpha_{i-1} - (\alpha_i + \alpha_{i+1})$ in $P_{n-1}$,
which sum to $\pi - \alpha_i - \alpha_{i+1}$;
it is similar for $j = i+2$.

@@

\begin{theorem}
Let $K$ be a weakly prime, alternating link whose diagram has no bigons.
Let $L$ be a link obtained from augmenting $K$.
Then $L$ is hyperbolic.
\end{theorem}

@@

By \prpref{p:tori_decomp}, $\toruscomp{L}$ can be
obtained by gluing two torihedra $\sT_T(L),\sT_B(L)$
with graphs $\Gamma_T(L),\Gamma_B(L)$.
We make them degenerate angled torihedra by assigning $\theta^*$'s
to edges of the bow-ties as in \figref{f:bowtie_angles},
and assign $\pi/2$ to all other edges.
It is easy to check that upon gluing,
each edge has sum of dihedral angles ($\theta^*$) equals $2\pi$.

@@

Furthermore, we can obtain an angle-splitting of $\sT_T(L)$
(and similarly $\sT_B(L)$) by modifying the angle-splitting
for $\sT_T(K)$;
this is shown in \figref{TODO}.

Now we have a decomposition of the two torihedra into
degenerate base-angled pyramids.
However, we need the pyramids to be non-degenerate,
so we first need to modify the angles and graph on the torihedra
to make all $\theta^*$ nonzero.

Consider a face $f$ of $\Gamma_T(L)$ that is not in a bow-tie.
Suppose the corresponding face $\bar{f}$ of $\Gamma$,
the projection graph (which is equal to $\Gamma_T(K)$),
had vertices $v_1,\ldots,v_n$ in counter-clockwise order.
We label the edges of $f$ by $e_{i,0}$, $e_{i,\pi}$, or $e_i$,
depending on whether the $\theta^*$ is $0$, $\pi$, or $\pi/2$ respectively,
with $i$ non-decreasing from 1 to $n$,
adjacent edges having the same $i$ if and only if they
belong to the same bow-tie.
For sake of concreteness,
suppose that if a vertex $v_i$ is right-augmented,
then the augmentation circle intersects $\bar{f}$
(everything is similar if it is left-augmented vertices' circles
that intersect $\bar{f}$).
In other words, locally, $f$ meets two of the edges of the bow-tie
corresponding to a right-augmented vertex $v_i$
(which would be labeled $e_{i,0}, e_{i,\pi}$ in counter-clockwise order),
but only meets one of the edges of the bow-tie corresponding to
a left-augmented vertex.

Suppose, after cyclically reindexing, $v_1,\ldots,v_k$
is a maximally contiguous subsequence of left-augmented vertices
of $G(K)$ around the face $\bar{f}$;
the edges around $f$ would start
$e_{1,0}, e_{1,\pi}, e_{2,0}, e_{2,\pi}, \ldots$.
We add new edges across $f$ as follows.
(See \figref{f:adding_edges}; the + and - signs will be explained later.)

@@

Next suppose there is only one such maximal contiguous subsequence.
If $k = 1$, we add an edge that goes across
$e_{1,0},e_{1,\pi},e_2$
(in the sense that the new edge separates the edges of $f$ into two sets
one of them being those three edges;
since $n\geq 3$, this edge is new).
%If $k = 2$, we add edge across $e_{1,0},e_{1,\pi}$
%and another edge across $e_{2,0},e_{2,\pi}$
%(these two edges do not form a bigon because we've ruled out $k=n$).
%(these two edges do not form a bigon because we've ruled out $k=n$).
If $k \geq 2$,
we add an edge across $e_{1,0},e_{1,\pi}$
and another edge across $e_{2,0},e_{2,\pi},e_{3,0},\ldots,e_{k,\pi}$
(these two edges do not form a bigon because we've ruled out $k=n$).
%(again these two edges do not form a bigon).

@@

This way we obtain a new graph $\Gamma_T(L)'$, which defines a
new torihedron $\sT_T(L)'$.
We make it angled using the angles from $\sT_T(L)$ for old edges,
and putting $\pi$ for all new edges.

\begin{figure}
\label{f:adding_edges}
\includegraphics[width=5cm]{more_pictures/one_edge.png}
%\includegraphics[width=5cm]{more_pictures/two_edge.png}
\includegraphics[width=5cm]{more_pictures/two_edge_many.png}
\end{figure}

Now we deform the $\theta^*$ based on \figref{f:adding_edges},
increasing/decreasing by some small $\veps' > 0$ if the edge
is labeled $+/-$.
Note some edges may be labeled twice, in which case we perform
both increasing/decreasing (e.g. if it is labeled $+$ and $-$,
the $\theta^*$ is not changed).
It is easy to see that the sum of $\theta^*$ around a vertex
remains unchanged.
Furthermore, all the edges with $\theta^*$ originally equal 0,
i.e. all $e_{i,0}$'s,
now have positive $\theta^*$,
(it receives only one label $+$,
because the other face it meets is a bow-tie).

Thus, we may now apply the feasible flow theorem in a similar manner,
and obtain an angle-splitting for $\sT_T(L)'$.
Alternatively, we can directly get an angle-splitting for $\sT_T(L)'$
using the angle-splitting for $\sT_T(L)$.
We reuse the $+/-$ assignments from \figref{f:adding_edges}.
For $e = e_{i,0}$, we increase $\vphi_{\vec{e}}, \vphi_{\cev{e}}$
by $\veps'/2$ each; for $e = e_{i,\pi}$,
we decrease them by $\veps'/2$ each.
For other edges, we increase/decrease $\vphi_{\vec{e}}$ by $\veps'$,
where $\vec{e}$ is the oriented edge corresponding to the side
on which the $+/-$ sign appears in \figref{f:adding_edges};
so for example, if an edge $e$ receives both $+$ and $-$,
then one of $\vphi_{\vec{e}},\vphi_{\cev{e}}$ increases
while the other decreases, thus $\theta_e^*$ remains constant.

Now we address $\sT_B(L)$.
In the gluing of $\sT_T(L)$ to $\sT_B(L)$,
non-bow-tie faces of $\Gamma_T(L)$ are identified with
non-bow-tie faces of $\Gamma_B(L)$.
Under this identification, we add the same edges to $\Gamma_B(L)$,
thus obtaining the new torihedron $\sT_B(L)'$ with graph
$\Gamma_B(L)'$.
We perform the same deformations of $\theta^*$'s (or $\vphi$'s).

We need to check that upon gluing $\sT_T(L)'$ to $\sT_B(L)'$,
the sum of dihedral angles around each edge is still $2\pi$.
This was clearly true before deforming, as the new edges of
$\Gamma_T(L)'$ only gets identified with the unique
corresponding edge of $\Gamma_B(L)'$, and they are both labeled
with $\theta^* = \pi$.
To see that the deformation does not change these sums,
note that in the identification of faces of $\Gamma_T(L)'$
to $\Gamma_B(L)'$,
an edge with $\theta^*=0$ is identified with an edge with
$\theta^*=\pi$,
so an increase in the former would be conterbalanced by
a decrease in the latter.
It is also easy to see this is the case for the other edges.
TODO hmm more detail?

Finally, for each face of $\Gamma_T(L)'$ that has more than three sides,
we arbitrarily decompose it into triangles
and apply \lemref{l:pyramid_decomp}
to obtain a triangulation of $sT_T(L)'$ into non-degenerate angled tetrahedra;
perform the corresponding decomposition for faces of
$\Gamma_B(L)'$ and obtain a triangulation of $\sT_B(L)'$
into non-degenerate angled tetrahedra.
Upon gluing, this gives an angle structure on the triangulation
of $\toruscomp{L}$
\end{proof}

@@
